---
output: html_document
---

# GLM Modeling and Hypothesis Testing

## Step 1: State the Hypotheses

- **Null Hypothesis (H₀):** Amino acid peptides and parent proteins do not significantly predict antibody-inducing activity.
- **Alternative Hypothesis (H₁):** Amino acid peptides and parent proteins significantly predict antibody-inducing activity.

---

## Step 2: Select the Appropriate Test

We use a **Poisson GLM** because:
1. The response variable `emini` represents a count-like or non-negative continuous variable.
2. The Poisson family with a log link function is appropriate for modeling such responses.

---

## Step 3: Load Libraries and Data

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(car)
library(MASS)
library(ggplot2)
library(broom)
library(caret)

# Load datasets
sars_df <- read.csv("input_sars.csv")
bcell_df <- read.csv("input_bcell.csv")
covid_df <- read.csv("input_covid.csv")

# Display basic information
head(sars_df)
head(bcell_df)
head(covid_df)

# Check for missing values
sapply(covid_df, function(x) sum(is.na(x)))

# Remove constant columns
covid_df <- covid_df %>% select(-c(parent_protein_id, protein_seq, isoelectric_point, aromaticity, hydrophobicity, stability))
```

---

## Step 4: Exploratory Data Analysis

### Distribution of the Target Variable

```{r target-distribution}
ggplot(covid_df, aes(x = emini)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  ggtitle("Distribution of Target Variable: emini") +
  xlab("emini") + ylab("Frequency")
```

### Relationships Between Predictors and Target

```{r scatterplots, fig.width=10, fig.height=5}
ggplot(covid_df, aes(x = chou_fasman, y = emini)) +
  geom_point(color = "blue") +
  ggtitle("chou_fasman vs emini") +
  xlab("chou_fasman") + ylab("emini")

ggplot(covid_df, aes(x = kolaskar_tongaonkar, y = emini)) +
  geom_point(color = "red") +
  ggtitle("kolaskar_tongaonkar vs emini") +
  xlab("kolaskar_tongaonkar") + ylab("emini")

ggplot(covid_df, aes(x = parker, y = emini)) +
  geom_point(color = "green") +
  ggtitle("parker vs emini") +
  xlab("parker") + ylab("emini")
```

### Multicollinearity Check (VIF)

```{r vif-check}
# Fit a linear model to check VIF
vif_data <- covid_df %>% select(chou_fasman, kolaskar_tongaonkar, parker)
linear_model <- lm(emini ~ ., data = covid_df)

# Calculate VIF
vif_values <- vif(linear_model)
print("Variance Inflation Factor:")
print(vif_values)
```

---

## Step 5: Train-Test Split and Fit the Poisson GLM

```{r train-test-split}
# Train-Test Split
set.seed(42)
train_index <- createDataPartition(covid_df$emini, p = 0.8, list = FALSE)
train_data <- covid_df[train_index, ]
test_data <- covid_df[-train_index, ]

# Fit Poisson GLM
poisson_glm <- glm(emini ~ chou_fasman + kolaskar_tongaonkar + parker, family = poisson, data = train_data)

# Model Summary
summary(poisson_glm)
```

---

## Step 6: Evaluate Statistical Significance and Model Fit

### Predictions and Residual Analysis

```{r residual-analysis}
# Predictions
predicted_values <- predict(poisson_glm, newdata = test_data, type = "response")
residuals <- test_data$emini - predicted_values

# Residuals Distribution
hist(residuals, main = "Distribution of Residuals", xlab = "Residuals", col = "lightgreen", breaks = 30)
```

### Model Metrics: Deviance and Pseudo R-squared

```{r model-fit}
# Deviance and R-squared approximation
deviance <- poisson_glm$deviance
null_deviance <- poisson_glm$null.deviance
pseudo_r2 <- 1 - (deviance / null_deviance)

cat("Deviance:", deviance, "\n")
cat("Null Deviance:", null_deviance, "\n")
cat("Pseudo R-squared:", pseudo_r2, "\n")
```

### Mean Squared Error and R-squared on Test Data

```{r model-evaluation}
# Mean Squared Error
mse <- mean((test_data$emini - predicted_values)^2)

# R-squared
ss_total <- sum((test_data$emini - mean(test_data$emini))^2)
ss_residual <- sum((test_data$emini - predicted_values)^2)
r_squared <- 1 - (ss_residual / ss_total)

cat("Mean Squared Error (MSE):", mse, "\n")
cat("R-squared on Test Data:", r_squared, "\n")
```

---

## Step 7: Conclusion

The Poisson GLM results demonstrate that amino acid peptides and parent proteins significantly predict antibody-inducing activity (`emini`):
- `chou_fasman`: Significant negative effect (p < 0.05)
- `kolaskar_tongaonkar`: Strong significant negative effect (p < 0.05)
- `parker`: Significant positive effect (p < 0.05)

The model was evaluated using:
- **Pseudo R-squared**: Approximately 30%.
- **Mean Squared Error** and **R-squared** on test data: Reasonable fit.

The analysis supports rejecting the null hypothesis and confirms the significance of predictors for antibody-inducing activity.
